---
title: "List of Myki cards"
output: html_notebook
date: 31/7/2018
---

## Introduction

 Code modified from the "code_to_get_started.R" script provided by the competition.
 
### Objective  
1.  Random sample of cards.
  + Proceed through each file and extract the cards and the frequency.

```{r setup}

#install.packages('data.table')
library(data.table)
library(tidyverse)
library(skimr)
library(gridExtra)
library(maps)

ScanOnFolderMaster <- '.././Samp_x/ScanOnTransaction'
ScanOffFolderMaster <- '.././Samp_x/ScanOffTransaction'

```

## Initial Data Inspection
Check out the data
```{r allFiles}
mySamp <- 0

ScanOnFolder <- sub("x", mySamp, ScanOnFolderMaster)
ScanOffFolder <- sub("x", mySamp, ScanOffFolderMaster)

#list the files
onFiles <- list.files(ScanOnFolder, recursive = TRUE, full.names = TRUE)
offFiles <- list.files(ScanOffFolder, recursive = TRUE, full.names = TRUE)

#how many
allFiles <- union(onFiles, offFiles)
cat("\nthere are", length(allFiles),'files')

```

```{r cards}
#------------------------------------
#read in a file and take a look
#------------------------------------
myFile <- onFiles[1]
cmd <- paste0("gzip -dc ", myFile)
cards <- fread(cmd, select = c(2, 4, 5)) # ,nrow=10000)

#these are the column names
colnames(cards) <- c('BusinessDate', 'CardID','CardType')

# dt %>% glimpse(width = 100)

cards <- cards %>% 
  mutate(BusinessDate = parse_date(BusinessDate, "%Y-%m-%d")) %>%   # "2015-07-02"
  glimpse(width = 100)

cards %>% 
  group_by(CardID, CardType) %>% 
  summarise(bus_date_min = min(BusinessDate),
            trans_freq = n()) %>%
  print()

```

OK, lets progress through each of the files and recursively build a list of CardIDs.  I'm assuming that `r 86011 * 316` will be too large to handle.  

```{r}
# 86011 * 316

# Initialise the receiving tibble.
cards_all <- tibble(CardID = vector("integer", 0),
                    CardType = vector("integer", 0),
                    bus_date_min = as.Date(NA),
                    trans_freq = vector("integer", 0))

for (fl in seq_along(onFiles)) {
  #  Import each file one by one and add to the list of CardIDs.
  
  # fl = 1
  cat(fl, ",")
  
  myFile <- onFiles[fl]
  cmd <- paste0("gzip -dc ", myFile)
  cards <- fread(cmd, select = c(2, 4, 5))
  
  #these are the column names
  colnames(cards) <- c('BusinessDate', 'CardID','CardType')
  
  cards <- cards %>% 
    mutate(BusinessDate = parse_date(BusinessDate, "%Y-%m-%d"))
  
  cards_all_new <- cards %>% 
    group_by(CardID, CardType) %>% 
    summarise(bus_date_min = min(BusinessDate),
              trans_freq = n()) %>%
    bind_rows(cards_all)
  
  cards_all <- cards_all_new %>% 
    group_by(CardID, CardType) %>% 
    summarise(bus_date_min = min(bus_date_min),
              trans_freq = sum(trans_freq, na.rm = TRUE))
  
}

cards_all


```

**Summary of cards**  
+ There are 1.5M cards.  
+ The card type ranges from 0 to 71.  
+ Dates range from Jul-15 to Jun-18.

```{r}
# skim(cards_all)

cards_all %>% summary()
```

```{r}

# Initialise the receiving tibble.
cards_all_off <- tibble(CardID = vector("integer", 0),
                    CardType = vector("integer", 0),
                    bus_date_min = as.Date(NA),
                    trans_freq = vector("integer", 0))

for (fl in seq_along(offFiles)) {
  #  Import each file one by one and add to the list of CardIDs.
  
  # fl = 1
  cat(fl, ",")
  
  myFile <- offFiles[fl]
  cmd <- paste0("gzip -dc ", myFile)
  cards <- fread(cmd, select = c(2, 4, 5))
  
  #these are the column names
  colnames(cards) <- c('BusinessDate', 'CardID','CardType')
  
  cards <- cards %>% 
    mutate(BusinessDate = parse_date(BusinessDate, "%Y-%m-%d"))
  
  cards_all_new <- cards %>% 
    group_by(CardID, CardType) %>% 
    summarise(bus_date_min = min(BusinessDate),
              trans_freq = n()) %>%
    bind_rows(cards_all_off)
  
  cards_all_off <- cards_all_new %>% 
    group_by(CardID, CardType) %>% 
    summarise(bus_date_min = min(bus_date_min),
              trans_freq = sum(trans_freq, na.rm = TRUE))
  
}

cards_all_off

```

```{r}
cards_both_summ <- cards_all %>% 
  rename(bus_date_on_min = bus_date_min,
         trans_on_freq = trans_freq) %>% 
  full_join(cards_all_off %>% rename(bus_date_off_min = bus_date_min,
                                     trans_off_freq = trans_freq),
            by = c("CardID", "CardType")) %>% 
  mutate(trans_both = coalesce(trans_on_freq, 0L) + coalesce(trans_off_freq, 0L)) %>% 
  glimpse()
```


```{r}
cards_both_summ %>% 
  ungroup() %>% 
  select(trans_on_freq, trans_off_freq) %>% 
  summarise(on_freq_sum = sum(trans_on_freq, na.rm = TRUE),
            on_freq_na = sum(is.na(trans_on_freq) | trans_on_freq == 0),
            off_freq_sum = sum(trans_off_freq, na.rm = TRUE),
            off_freq_na = sum(is.na(trans_off_freq) | trans_off_freq == 0))
```

```{r}
# write_rds(cards_both_summ, "cards_both_summ.rds")

cards_both_summ <- read_rds("cards_both_summ.rds")

```

```{r}
cards_both_summ %>% 
  group_by(CardID) %>% 
  mutate(CardType_num = n()) %>% 
  ungroup() %>% 
  arrange(-CardType_num, CardID)
```

```{r}
cards_both_summ <- cards_both_summ %>% 
  ungroup() %>% 
  group_by(CardID) %>% 
  summarise(bus_date_on_min = min(bus_date_on_min, na.rm = TRUE),
            trans_on_freq = sum(trans_on_freq, na.rm = TRUE),
            bus_date_off_min = min(bus_date_off_min, na.rm = TRUE),
            trans_off_freq = sum(trans_off_freq, na.rm = TRUE),
            trans_both = sum(trans_both, na.rm = TRUE))
```


