---
title: "Myki data EDA"
output: html_notebook
date: 24/7/2018
---

## Introduction

 Code modified from the "code_to_get_started.R" script provided by the competition

```{r setup}

#install.packages('data.table')
library(data.table)
library(tidyverse)
library(skimr)
library(gridExtra)
library(maps)

ScanOnFolderMaster <- '.././Samp_x/ScanOnTransaction'
ScanOffFolderMaster <- '.././Samp_x/ScanOffTransaction'

```

## Initial Data Inspection
Check out the data
```{r allFiles}
mySamp <- 0

ScanOnFolder <- sub("x", mySamp, ScanOnFolderMaster)
ScanOffFolder <- sub("x", mySamp, ScanOffFolderMaster)

#list the files
onFiles <- list.files(ScanOnFolder, recursive = TRUE, full.names = TRUE)
offFiles <- list.files(ScanOffFolder, recursive = TRUE, full.names = TRUE)

#how many
allFiles <- union(onFiles, offFiles)
cat("\nthere are", length(allFiles),'files')

```


```{r dt}
#------------------------------------
#read in a file and take a look
#------------------------------------
myFile <- onFiles[1]
cmd <- paste0("gzip -dc ", myFile)
dt <- fread(cmd) # ,nrow=10000)

#these are the column names
colnames(dt) <- c('Mode','BusinessDate','DateTime','CardID','CardType','VehicleID','ParentRoute','RouteID','StopID')

# dt %>% glimpse(width = 100)

dt <- dt %>% 
  mutate(BusinessDate = parse_date(BusinessDate, "%Y-%m-%d"),
         DateTime = parse_datetime(DateTime, "%Y-%m-%d %H:%M:%S")) %>%   # "2015-07-02 13:08:13"
  glimpse(width = 100)

```



```{r}
skim(dt)
```

Quick plot of each of the numeric fields.  Interestingly, this file has only 4 days of data.
```{r, warning=FALSE, fig.height=20, fig.width=7}

dt_num_lst <- dt %>% 
  select_if(is_numeric) %>% 
  colnames()

dt_plot_lst <- vector("list", length(dt_num_lst))

for (i in seq_along(dt_num_lst)) {
  # print a histogram

  col_name = dt_num_lst[i]
  
  dt_plot <-
    dt %>%
    ggplot(aes_string(x = col_name)) +
    geom_histogram(bins = 30) +
    labs(title = str_c("Distribution of ", col_name, collapse = ""))

  dt_plot_lst[[i]] <- dt_plot
  
}

grid.arrange(grobs = dt_plot_lst, ncol = 1)


```

### Questions from the plots  

1. Is there any significance with the stop numbers?  
  + Maybe could a map plot with colour coding the stop numbers.  
2. Why is DateTime so lumpy?
  + Suspect that it is due to the bins that mean that there is a pattern to whether the bin includes multiple peak periods.  
3. Latitude and longitude plots look reasonable.  


```{r}
aus <- map_data("world2", region = "au")

ggplot(aus %>% filter(between(long, 144.4, 145.6),
                      between(lat, -38.5, -37.0)), aes(long, lat, group = group)) +
  geom_polygon() +
  coord_quickmap()

# aus %>% 
#   count(group, subregion, sort = TRUE)
# 
# ggmap::geocode("AU")
```

```{r, warning=FALSE, fig.height=10, fig.width=10}
library(ggmap)
# citation('ggmap') 

map_melb <- get_map(location = c(144.4, -38.3, 145.6, -37.5), zoom = 9) %>%
ggmap(alpha = 0.5)

# plot(bt$lat,bt$long,col=factor(bt$CardID),pch=19,main="busy travellers")

stops <- bt %>% 
  select(StopID, lat, long) %>% 
  distinct() %>% 
  glimpse()

map_melb +
  geom_point(data = stops, aes(x = lat, y = long), size = 200, colour = "red")

```

