---
title: "Collect transaction samples"
output: html_notebook
date: 1/8/2018
---

## Introduction

The purpose of this notebook is to collect sample Myki transaction.

A prerequisite to running this notebook is the data of Myki cards.

```{r setup}

#install.packages('data.table')
library(data.table)
library(tidyverse)
library(skimr)
library(gridExtra)
library(maps)

ScanOnFolderMaster <- '.././Samp_x/ScanOnTransaction'
ScanOffFolderMaster <- '.././Samp_x/ScanOffTransaction'

# Set the proportion of cards to analyse
card_samp_sh <- 0.01

trans_col_names <- c('Mode','BusinessDate','DateTime','CardID','CardType','VehicleID','ParentRoute','RouteID','StopID')

```


```{r}
# write_rds(cards_both_summ, "cards_both_summ.rds")

cards_both_summ <- read_rds("cards_both_summ.rds")

cards_both_summ <- cards_both_summ %>% 
  ungroup() %>% 
  group_by(CardID) %>% 
  summarise(bus_date_on_min = min(bus_date_on_min, na.rm = TRUE),
            trans_on_freq = sum(trans_on_freq, na.rm = TRUE),
            bus_date_off_min = min(bus_date_off_min, na.rm = TRUE),
            trans_off_freq = sum(trans_off_freq, na.rm = TRUE),
            trans_both = sum(trans_both, na.rm = TRUE)) %>% 
  ungroup()

```

### Pick sample

```{r}

card_samp <- sample_frac(cards_both_summ, size = card_samp_sh, weight = trans_on_freq) %>% 
  print()

```

```{r}
cards_both_summ %>% summary()
```

I  deliberately weighted the sample by the number of touch-on transactions, as opposed to touch-on and touch-off transactions.

I may want to revisit this assumption - are we interested in tracking individual cards or transactions?

```{r}
card_samp %>% skim(.)
```

## Pick out the sample transactions

```{r}

# Initialise the receiving tibble.
card_trans_lst <- vector("list", length(onFiles))

for (fl in seq_along(onFiles)) {
  #  Import each file one by one and add to the list of CardIDs.
  
  # fl = 1
  cat(fl, ",")
  
  myFile <- onFiles[fl]
  cmd <- paste0("gzip -dc ", myFile)
  card_trans <- fread(cmd)
  
  #these are the column names
  colnames(card_trans) <- trans_col_names
  
  card_trans <- card_trans %>% 
    inner_join(card_samp %>% select(CardID),
               by = "CardID")
  
  card_trans <- card_trans %>% 
    mutate(BusinessDate = parse_date(BusinessDate, "%Y-%m-%d"),
           DateTime = parse_datetime(DateTime, "%Y-%m-%d %H:%M:%S"))   # "2015-07-02 13:08:13"
  
  card_trans_lst[[fl]] <- card_trans
  
}

trans_on_samp <- bind_rows(card_trans_lst)


```

```{r}

# Initialise the receiving tibble.
card_trans_lst <- vector("list", length(offFiles))

for (fl in seq_along(offFiles)) {
  #  Import each file one by one and add to the list of CardIDs.
  
  # fl = 1
  cat(fl, ",")
  
  myFile <- offFiles[fl]
  cmd <- paste0("gzip -dc ", myFile)
  card_trans <- fread(cmd)
  
  #these are the column names
  colnames(card_trans) <- trans_col_names
  
  card_trans <- card_trans %>% 
    inner_join(card_samp %>% select(CardID),
               by = "CardID")
  
  # card_trans <- card_trans %>% 
  #   mutate(BusinessDate = parse_date(BusinessDate, "%Y-%m-%d"),
  #          DateTime = parse_datetime(DateTime, "%Y-%m-%d %H:%M:%S"))   # "2015-07-02 13:08:13"
  
  card_trans_lst[[fl]] <- card_trans
  
}

trans_off_samp <- bind_rows(card_trans_lst) %>% 
    mutate(BusinessDate = parse_date(BusinessDate, "%Y-%m-%d"),
           DateTime = parse_datetime(DateTime, "%Y-%m-%d %H:%M:%S"))   # "2015-07-02 13:08:13"


```

```{r}
write_rds(trans_on_samp, "trans_on_samp.rds")

write_rds(trans_off_samp, "trans_off_samp.rds")

```


